---
title: "Exploratory Graph Analysis"
author: "Daniel van der Meer, Fridtjof Petersen, Jonas Petter"
date: "4/26/2022"
output: 
  html_document:
    toc: true
    toc_float: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
```

## Introduction to exploratory graph analysis

This tutorial illustrates how to apply exploratory graph analysis (EGA) using R. EGA was proposed by Golino and Epskamp (2017) as an alternative to common methods of exploratory factor analysis (EFA). EGA is based on the insight that latent variables are reflected as clusters in network models. Broadly speaking, EGA consists of three steps: 

1) Estimate the correlation matrix.

2) Construct a network graph from the correlation matrix.

3) Identify clusters in the network graph.

This tutorial will illustrate how to apply these steps in practice and then draw a short comparison to other methods of EFA.

## The dataset 
```{r}
library(lavaan)
data <- HolzingerSwineford1939[7:15]
```

In this example, we will use the data set from Holzinger and Swineford (1939), which comes with the R package *lavaan*. The researchers measured 300 students on their cognitive abilities, using a 9-item-questionnaire. These 9 items are assumed to correspond to visual, verbal and speed abilities. This 3 factor structure is indeed known to achieve a high fit with the data set. Thus, we hope to recover this 3 factor structure using EGA.

## Applying exploratory graph analysis step-by-step

The following code is adapted from Golino and Epskamp (2017).

Since EGA investigates clusters in network models, we need to create a network model of the data. For this we first need to get the correlation matrix of the 9 items. This is easily achieved with the *cor_auto* function from *qgraph*.

```{r}
library(qgraph)
correlation_matrix <- cor_auto(data)
```

Then, we can estimate the Gaussian graphical model, a network model, using the graphical lasso (GLASSO) method. For this, we can use the *EBICglasso* function from *qgraph*.

```{r}
glasso <- EBICglasso(S=correlation_matrix, n = nrow(data))
```

Let us take a look at our estimated network model, using the qgraph function.

```{r}
networkGraph <- qgraph(glasso, layout = "spring", vsize = 6)
```

We can already see that the clusters in this network model correspond to our assumed factor structure. But to verify this mathematically, we need to take a few more steps. To obtain clusters from a network, EGA uses the walktrap algorithm. This algorithm is implemented in the *igraph* package. To use it, we need to convert our network model to the igraph notation and then use the *walktrap.community* function.

```{r}
library(igraph)
igraph <- as.igraph(networkGraph)
walktrapResults <- walktrap.community(igraph)
walktrapResults
```

The results tell us that there were indeed 3 clusters. This corresponds to a 3 factor structure, as expected. Factor 1 loads onto items 7, 8 and 9. Factor 2 loads onto items 1, 2 and 3. Factor 3 loads onto items 4, 5 and 6. This is in line with the proposed factor structure of visual, verbal and speed abilities. Finally, let us visualize the clusters using *qgraph*.

```{r}
qgraph(glasso, layout = "spring", vsize = 6, groups = as.factor(walktrapResults$membership))
```


## Packages for exploratory graph analysis

Fortunately, it is not necessary to go through all of the steps of EGA manually. Golino et al. have authored the R package *EGAnet*, which implements the EGA framework in a single function. This function gives us the same output as in the previous section, but with much less code. 

```{r}
library(EGAnet)
EGAnet::EGA(data, model = "glasso", plot.EGA = TRUE)
```


## Comparison with other methods

Alternatively, we can perform a parallel analysis to determine the number of factors. For this we can use the *parallel* function from the *nFactors* package.

```{r}
library(nFactors)
eigenvalues <- eigen(correlation_matrix)
parallelAnalysis <- parallel(subject = nrow(data), var = ncol(data), rep = 100, cent = .05)
nscree <- nScree(eigenvalues$values, aparallel = parallelAnalysis$eigen$qevpea)
plotnScree(nscree)
```

The Kaiser criterion recommends to drop all factors with an eigenvalue < 1. From this analysis we would thus also conclude that a 3-factor structure is appropriate for the data. The parallel analysis comes to the same conclusion. 

Let's now take a look at the factor loadings, using the *fa* function from the *psych* package.

```{r}
library(psych)
efa <- fa(data, nfactors = 3, rotate = "varimax")
efa$loadings
```

Looking at the items with the highest loadings on each factor, we can recover the originally proposed structure. Items 4, 5, and 6, load strongest on factor 1. Items 7, 8 and 9 load strongest on Factor 2. Only for factor 3 the conclusion is less clear than from the EGA. While items 1, 2, and 3 do have the highest loadings, item 9 also has a reasonably high loading on this factor.

## References

Golino, H. F., & Epskamp, S. (2017). Exploratory graph analysis: A new approach for estimating the number of dimensions in             psychological research. *PLOS ONE, 12*(6), e0174035. doi:10.1371/journal.pone.0174035

Golino, H., & Christensen, A. P. (2021). EGAnet: Exploratory Graph Analysis – A framework for estimating the number
  of dimensions in multivariate data using network psychometrics. R package version 0.9.9.
  
Holzinger, K. J., & Swineford, F. (1939). A study in factor analysis: The stability of a bi-factor solution. *Supplementary          educational monographs*.
  

